```{r}
library(dplyr)
library(tidyverse)
library(readxl)
```
**loading prior data**
Mut_prev is Mut+mix/N
```{r}
published_historical <- read_excel("~/Rosenthal Lab/Tororolong/Plots /Feb 10 plot/published_historical.xlsx")

published_historical <- published_historical %>% 
  filter(Site == "Tororo") %>%
  mutate("Mut_N" = Mix + Mut) 
  
```



**prior**
*The beta distribution for Tororo 2016, a very "informative" prior*
dbeta(x_vals, shape 1 = a = Mut, shape 2 = b = WT)
our measure, pi, is distributed dbeta(x_vals, shape 1 = a = 32, shape 2 = b = 32)

pi ~ beta(32,32)
```{r}
library(ggplot2)

# Generate Y_vals
X_vals <- seq(0, 1, by = 0.01)

# Calculate the density values using dbeta
densities_informative <- dbeta(X_vals, 32, 32)
densities_uninformative <- dbeta(X_vals, 8, 8)

# Plotting the beta density
ggplot() +
  geom_line(aes(x = X_vals, y = densities_informative, color = "Informative"), size = 0.75) +
  geom_line(aes(x = X_vals, y = densities_uninformative, color = "Uninformative"), size = 0.75) +
  labs(title = "Density of PfCRTK76T, informative vs uninformative prior information",
       x = "Prevalence",
       y = "Density") +
  scale_color_manual(values = c("Informative" = "lightblue", "Uninformative" = "darkgreen"), 
                     labels = c("Informative Prior", "Uninformative Prior")) +
  theme_minimal()


```


**likelihood distribution**
Y ∼ Binomial(N, pi)
Y ∼ Binomial(27, .48)
```{r}
library(ggplot2)

# Generate X_vals representing the number of successes
X_vals_binom <- 0:27

# Calculate the binomial likelihood function using dbinom
likelihood <- dbinom(X_vals_binom, size = 27, prob = 0.48)

# Create a data frame for plotting
likelihood_df <- data.frame(X_vals_binom, likelihood)

# Plot the binomial likelihood
ggplot(likelihood_df, aes(x = X_vals_binom, y = likelihood)) +
  geom_line(color = "coral") +
  labs(title = "Binomial Likelihood Function (data I generated)",
       x = "Mut_N",
       y = "Likelihood") + theme_minimal()



```
As we can see, the probability is highest around 13



**prior and likelihood together**
described as a conjugate beta distribution 

_new data_
we see y = 13 infections = Mut_N
n = 27 infections = N

Pi = .48

_prior_
a = Mut_N = 32
b = WT_N = 32


Beta(a + y, b + n − y)
```{r}
post_X <- seq(0,1, .01)

posterior <- dbeta(post_X, shape1 = 32 + 13, shape2 = 32 + 27 - 13)

library(ggplot2)

ggplot() +
  geom_line(aes(x = post_X, y = posterior), color = "purple") +
  labs(title = "Posterior Prevalence of PfCRT K76T",
       x = "Prevalence",
       y = "Posterior Likelihood") +
  theme_minimal() 


```
```{r}
qbeta(.5, shape1 = 32 + 13, shape2 = 32 + 27 - 13)
```

**making the percolatR**
**specifying posterior distribution scaled by N**
Because we are specifying our binomial dist as our 
```{r}

percolateR <- percolateR <- function(Mut_N_Prior, WT_N_Prior, Mut_N_New, N_New){
  
  a_post <- Mut_N_Prior + Mut_N_New #shape parameter in posterior 
  b_post <- WT_N_Prior + N_New - Mut_N_New
  
  posterior <- round(qbeta(c(0.5, 0.025, 0.975), a_post, b_post), 3)
  
  return(tibble::tibble(
    "Posterior Prevalence" = posterior[1],
    "Posterior LB" = posterior[2],
    "Posterior UB" = posterior[3]
  ))

}
```

```{r}
#testing the percolator on my 2012 vs other 2012 data 

percolateR(Mut_N_Prior = 12, WT_N_Prior = 3, Mut_N_New = 17, N_New = 18)

#it works
```

**expanding the percolatR out for df's**
```{r}
library(readr)
filt_bayes <- read_csv("~/Rosenthal Lab/Tororolong/Plots /Feb 10 plot/filt_bayes.csv")


df_prior_e<- published_historical %>% select(Marker, Year, WT, Mut_N) 

df_likelihood <- filt_bayes


df_likelihood

```

```{r}
#reformatting the table for prior 
data_long <- na.omit(pivot_longer(df_likelihood, 
                          cols = -Locus,
                          names_to = c(".value", "Date"),
                          names_pattern = "(Mut_N)_(\\d+)"))

data_long2 <- na.omit(pivot_longer(df_likelihood, 
                          cols = -Locus,
                          names_to = c(".value", "Date"),
                          names_pattern = "(N) (\\d+)"))


true_likelihood <- left_join(data_long, data_long2, by = c("Locus", "Date"))

```

renaming the columns
```{r}
df_prior <- df_prior_e %>%
  rename("WT_N_Prior" = WT,
         "Mut_N_Prior" = Mut_N)

df_likelihood <- true_likelihood %>%
  rename("N_New" = N,
         "Mut_N_New" = Mut_N,
         "Year" = Date)



```

```{r}

#this function just joins the dfs by date filtering and by via marker number
#for example if I wanted 
percolateR_0.1 <- function(df_prior, df_likelihood, marker_number) {
  
  pripost <- left_join(
    df_prior %>% filter(grepl(marker_number, Marker)), 
    df_likelihood %>% filter(grepl(marker_number, Locus)),
    by = c("Year")
  )
  
  return(na.omit(pripost))
}

percolateR_0.1(df_prior, df_likelihood, 76)



```
```{r}

#this function just joins the dfs by date filtering and by via marker number
#for example if I wanted 
percolateR_0.2 <- function(df_prior, df_likelihood, marker_number) {
  
  #combining the df's based on marker number and year, all years represented in both the posterior and the likelihood dataset 
  pripost <- left_join(
    df_prior %>% filter(grepl(marker_number, Marker)), 
    df_likelihood %>% filter(grepl(marker_number, Locus)),
    by = c("Year")
  )
  
  #vectorizing the Ns
  Mut_N_Prior <- pripost$Mut_N_Prior
  WT_N_Prior <- pripost$WT_N_Prior
  
  N_New <- pripost$N_New
  Mut_N_New <- pripost$N_New
  
  
  #for every thing in the list do the bayesian tabulation
  for (i in 1:length(Mut_N_Prior, WT_N_Prior, N_New, Mut_N_New)) { 
    
  a_post[i] <- Mut_N_Prior[i] + Mut_N_New[i] #shape parameter in posterior 
  b_post[i] <- WT_N_Prior[i] + N_New[i] - Mut_N_New[i]
  
  
  posterior <- round(qbeta(c(0.5, 0.025, 0.975), a_post[i], b_post[i]), 3)
  
  
  }
  
  return(na.omit(pripost$posterior))
}

percolateR_0.2(df_prior, df_likelihood, 540)

```

```{r}
percolateR_0.2 <- function(df_prior, df_likelihood, marker_number) {
  
  # Combine the data frames based on marker number
  pripost <- left_join(
    df_prior %>% filter(grepl(marker_number, Marker)), 
    df_likelihood %>% filter(grepl(marker_number, Locus)),
    by = "Year"
  )
  
  # Vectorize the variables
  Mut_N_Prior <- pripost$Mut_N_Prior
  WT_N_Prior <- pripost$WT_N_Prior
  N_New <- pripost$N_New
  Mut_N_New <- pripost$Mut_N_New
  
  # Initialize vectors to store posterior values
  posterior_prevalence <- numeric(length(Mut_N_Prior))
  posterior_lb <- numeric(length(Mut_N_Prior))
  posterior_ub <- numeric(length(Mut_N_Prior))
  
  # Calculate posterior values for each row
  for (i in 1:length(Mut_N_Prior)) { 
    a_post <- Mut_N_Prior[i] + Mut_N_New[i] # Shape parameter in posterior 
    b_post <- WT_N_Prior[i] + N_New[i] - Mut_N_New[i]
    posterior <- round(qbeta(c(0.5, 0.025, 0.975), a_post, b_post), 3)
    
    # Store posterior values in vectors
    posterior_prevalence[i] <- posterior[1]
    posterior_lb[i] <- posterior[2]
    posterior_ub[i] <- posterior[3]
  }
  
  # Add posterior values to the pripost data frame
  pripost <- bind_cols(pripost, 
                       "Posterior Prevalence" = posterior_prevalence,
                       "Posterior LB" = posterior_lb,
                       "Posterior UB" = posterior_ub)
  
  # Return the updated data frame
  return(na.omit(pripost))
}

```

```{r}
 dhfr51 <- percolateR_0.2(df_prior, df_likelihood, 51)

head(dhfr51) 



```
**
```{r}
percolateR(Mut_N_Prior = 0, WT_N_Prior = 0, Mut_N_New = 46.5, N_New = 48.5)

percolateR_0.5(df_prior, df_likelihood, 51)
```
```{r}
PfMDR1_86 <- percolateR_0.2(df_prior, df_likelihood, 86)
PfMDR1_148 <- percolateR_0.2(df_prior, df_likelihood, 148)
PfMDR1_1246 <- percolateR_0.2(df_prior, df_likelihood, 1246)

PfMDR1_86$Year <- as.numeric(PfMDR1_86$Year)
PfMDR1_148$Year <- as.numeric(PfMDR1_148$Year)
PfMDR1_1246$Year <- as.numeric(PfMDR1_1246$Year)



# Convert Year to numeric
PfMDR1$Year <- as.numeric(PfMDR1$Year)

# Plotting
library(ggplot2)




```
```{r}
percolateR_0.3 <- function(df_prior, df_likelihood, marker_number) {
  
  # Combine the data frames based on marker number and year
  pripost <- left_join(
    df_prior %>% filter(grepl(marker_number, Marker)), 
    df_likelihood %>% filter(grepl(marker_number, Locus)),
    by = "Year"
  )
  
  # Vectorize the variables
  Mut_N_Prior <- pripost$Mut_N_Prior
  WT_N_Prior <- pripost$WT_N_Prior
  N_New <- pripost$N_New
  Mut_N_New <- ifelse(is.na(pripost$Mut_N_New), pripost$Mut_N_New, pripost$N_New - pripost$Mut_N_New)
  
  # Calculate posterior values for each row
  posterior_prevalence <- numeric(length(Mut_N_Prior))
  posterior_lb <- numeric(length(Mut_N_Prior))
  posterior_ub <- numeric(length(Mut_N_Prior))
  
  for (i in 1:length(Mut_N_Prior)) { 
    a_post <- Mut_N_Prior[i] + Mut_N_New[i] # Shape parameter in posterior 
    b_post <- WT_N_Prior[i] + N_New[i] - Mut_N_New[i]
    posterior <- round(qbeta(c(0.5, 0.025, 0.975), a_post, b_post), 3)
    
    # Store posterior values in vectors
    posterior_prevalence[i] <- posterior[1]
    posterior_lb[i] <- posterior[2]
    posterior_ub[i] <- posterior[3]
  }
  
  # Add posterior values to the pripost data frame
  pripost <- bind_cols(pripost, 
                       "Posterior Prevalence" = posterior_prevalence,
                       "Posterior LB" = posterior_lb,
                       "Posterior UB" = posterior_ub)
  
  # Return the updated data frame
  return(na.omit(pripost))
}


```

```{r}
 CRTK76T<- percolateR_0.3(df_prior, df_likelihood, 76)

head(CRTK76T)
```


```{r}
percolateR_0.4 <- function(df_prior, df_likelihood, marker_number) {
  # Subset df_prior and df_likelihood to contain only the specified marker number
  df_prior <- df_prior %>% filter(grepl(marker_number, Marker))
  df_likelihood <- df_likelihood %>% filter(grepl(marker_number, Locus))

  # Initialize vectors to store posterior values
  posterior_prevalence <- numeric()
  posterior_lb <- numeric()
  posterior_ub <- numeric()
  corresponding_year <- logical()

  # Loop through each year in df_likelihood
  for (i in 1:nrow(df_likelihood)) {
    year <- df_likelihood$Year[i]

    # Check if there's a corresponding year in df_prior
    if (year %in% df_prior$Year) {
      # Combine the data frames based on marker number and year
      pripost <- left_join(
        df_prior %>% filter(Year == year), 
        df_likelihood %>% filter(Year == year),
        by = "Year"
      )

      # Vectorize the variables
      Mut_N_Prior <- pripost$Mut_N_Prior
      WT_N_Prior <- pripost$WT_N_Prior
      Mut_N_New <- pripost$Mut_N_New

      # Calculate posterior values for the current year
      a_post <- Mut_N_Prior + Mut_N_New
      b_post <- WT_N_Prior + pripost$N_New - Mut_N_New
      posterior <- round(qbeta(c(0.5, 0.025, 0.975), a_post, b_post), 3)

      # Store posterior values in vectors
      posterior_prevalence <- c(posterior_prevalence, posterior[1])
      posterior_lb <- c(posterior_lb, posterior[2])
      posterior_ub <- c(posterior_ub, posterior[3])
      corresponding_year <- c(corresponding_year, TRUE)
    } else {
      # No corresponding year in df_prior, calculate Mut_N_New/N_New proportions
      Mut_N_New_prop <- df_likelihood$Mut_N_New[i] / df_likelihood$N_New[i]

      # Calculate confidence interval based on proportions using prop.test
      prop_test <- prop.test(df_likelihood$Mut_N_New[i], df_likelihood$N_New[i], conf.level = 0.70)

      # Store posterior values in vectors
      posterior_prevalence <- c(posterior_prevalence, Mut_N_New_prop)
      posterior_lb <- c(posterior_lb, prop_test$conf.int[1])
      posterior_ub <- c(posterior_ub, prop_test$conf.int[2])
      corresponding_year <- c(corresponding_year, FALSE)
    }
  }

  # Create a data frame with the calculated values
  pripost <- data.frame(Year = df_likelihood$Year,
                        Locus = df_likelihood$Locus,
                        Mut_N_Prior = pripost$Mut_N_Prior,
                        WT_N_Prior = pripost$WT_N_Prior,
                        Mut_N_New = df_likelihood$Mut_N_New,
                        N_New = df_likelihood$N_New,
                        Corresponding_Year = corresponding_year,
                        Posterior_Prevalence = posterior_prevalence,
                        Posterior_LB = posterior_lb,
                        Posterior_UB = posterior_ub)

  # Return the updated data frame
  return(pripost)
}


```

```{r}
 CRTK76T<- percolateR_0.4(df_prior, df_likelihood, 76)
```

```{r}


CRTK76T$Year <- as.numeric(CRTK76T$Year)

ggplot(CRTK76T, aes(x = Year, y = `Posterior_Prevalence`)) +
  geom_line(color = "blue") +
  geom_ribbon(aes(ymin = `Posterior_LB`, ymax = `Posterior_UB`), fill = "lightblue", alpha = 0.5) +
  labs(x = "Year", y = "Posterior Prevalence",
       title = "Posterior Prevalence PfCRTK76T",
       subtitle = "Blue line represents posterior prevalence, light blue band represents 95% credible interval") +
  theme_minimal()
```
**using a beta distribution non informative when there is no corresponding year**
**fuck man**
```{r}

percolateR_0.5 <- function(df_prior, df_likelihood, marker_number) {
  # Subset df_prior and df_likelihood to contain only the specified marker number
  df_prior <- df_prior %>% filter(grepl(marker_number, Marker))
  df_likelihood <- df_likelihood %>% filter(grepl(marker_number, Locus))

  # Initialize vectors to store posterior values
  posterior_prevalence <- numeric()
  posterior_lb <- numeric()
  posterior_ub <- numeric()
  corresponding_year <- logical()
  
  # Initialize Mut_N_Prior and WT_N_Prior as NA vectors
  Mut_N_Prior <- rep(NA, nrow(df_likelihood))
  WT_N_Prior <- rep(NA, nrow(df_likelihood))

  # Loop through each year in df_likelihood
  for (i in 1:nrow(df_likelihood)) {
    year <- df_likelihood$Year[i]

    # Check if there's a corresponding year in df_prior
    if (year %in% df_prior$Year) {
      # Combine the data frames based on marker number and year
      pripost <- left_join(
        df_prior %>% filter(Year == year), 
        df_likelihood %>% filter(Year == year),
        by = "Year"
      )

      # Assign Mut_N_Prior and WT_N_Prior for the current year
      Mut_N_Prior[i] <- pripost$Mut_N_Prior
      WT_N_Prior[i] <- pripost$WT_N_Prior
    } else {
      # There's no corresponding year in df_prior
      # Use Jeffrey's prior for Mut_N_Prior and WT_N_Prior
      Mut_N_Prior[i] <- .5
      WT_N_Prior[i] <- .5
    }

    # Calculate posterior values for the current year
    a_post <- Mut_N_Prior[i] + df_likelihood$Mut_N_New[i]
    b_post <- WT_N_Prior[i] + df_likelihood$N_New[i] - df_likelihood$Mut_N_New[i]
    posterior <- round(qbeta(c(0.5, 0.025, 0.975), a_post, b_post), 3)

    # Store posterior values in vectors
    posterior_prevalence <- c(posterior_prevalence, posterior[1])
    posterior_lb <- c(posterior_lb, posterior[2])
    posterior_ub <- c(posterior_ub, posterior[3])
    corresponding_year <- c(corresponding_year, year %in% df_prior$Year)
  }

  # Create a data frame with the calculated values
  pripost <- data.frame(
    Year = df_likelihood$Year,
    Locus = df_likelihood$Locus,
    Mut_N_Prior = Mut_N_Prior,
    WT_N_Prior = WT_N_Prior,
    Mut_N_New = df_likelihood$Mut_N_New,
    N_New = df_likelihood$N_New,
    Corresponding_Year = corresponding_year,
    Posterior_Prevalence = posterior_prevalence,
    Posterior_LB = posterior_lb,
    Posterior_UB = posterior_ub
  )

  # Return the updated data frame
  return(pripost)
}



```
**the fifth iteration of the percolateR**
```{r}
CRTK76T.5 <- percolateR_0.5(df_prior = df_prior, df_likelihood = df_likelihood, 76)

CRTK76T.5$Year <- as.numeric(CRTK76T.5$Year)

head(CRTK76T.5)

kable(CRTK76T.5, "html") %>%
  kable_styling(full_width = FALSE) %>%
  row_spec(0, bold = TRUE) 
library(ggplot2)
```


```{r}


# Convert Year column to numeric
CRTK76T.5$Year <- as.numeric(CRTK76T.5$Year)


# Plotting
ggplot(CRTK76T.5) +
  geom_line(aes(x = Year, y = Posterior_Prevalence, color = "CRT-K76T")) +
  geom_ribbon(aes(x = Year, ymin = Posterior_LB, ymax = Posterior_UB, fill = "95% Credible Interval"), fill = "lightblue", alpha = 0.5) +
  labs(x = "Year", y = "Posterior Prevalence",
       title = "Posterior Prevalence of PfCRT Polymorphisms Over Time",
       subtitle = "Shade is 95% credible interval") +
  scale_color_manual(name = "Locus", values = c("CRT-K76T" = "darkblue")) +
  scale_fill_identity(name = "Legend", guide = "legend", labels = c("95% Credible Interval")) +
  theme_minimal()


```

making the table for crt
```{r}
library(kableExtra)

allMDR.5 <- rbind(PfMDR_86.5, PfMDR_184.5, PfMDR_1246.5)

kable(allMDR.5, "html") %>%
  kable_styling(full_width = FALSE) %>%
  row_spec(0, bold = TRUE) 
```

```{r}

# Calculate posterior prevalence for different scenarios
PfMDR_86.5 <- percolateR_0.5(df_prior, df_likelihood, 86)
PfMDR_184.5 <- percolateR_0.5(df_prior, df_likelihood, 184)
PfMDR_1246.5 <- percolateR_0.5(df_prior, df_likelihood, 1246)

# Convert Year column to numeric
PfMDR_86.5$Year <- as.numeric(PfMDR_86.5$Year)
PfMDR_184.5$Year <- as.numeric(PfMDR_184.5$Year)
PfMDR_1246.5$Year <- as.numeric(PfMDR_1246.5$Year)


```

```{r}
ggplot() +
  geom_line(data = PfMDR_86.5, aes(x = Year, y = Posterior_Prevalence, color = "MDR1-N86Y")) +
  geom_ribbon(data = PfMDR_86.5, aes(x = Year, ymin = Posterior_LB, ymax = Posterior_UB, fill = "MDR1-N86Y"), fill = "coral", alpha = 0.5) +
  geom_line(data = PfMDR_184.5, aes(x = Year, y = Posterior_Prevalence, color = "MDR1-Y184F")) +
  geom_ribbon(data = PfMDR_184.5, aes(x = Year, ymin = Posterior_LB, ymax = Posterior_UB, fill = "MDR1-Y184F"), fill = "lightgreen", alpha = 0.5) +
  geom_line(data = PfMDR_1246.5, aes(x = Year, y = Posterior_Prevalence, color = "MDR1-D1246Y")) +
  geom_ribbon(data = PfMDR_1246.5, aes(x = Year, ymin = Posterior_LB, ymax = Posterior_UB, fill = "MDR1-D1246Y"), fill = "lightblue", alpha = 0.5) +
  scale_color_manual(name = "Locus", values = c("MDR1-D1246Y" = "darkblue", "MDR1-Y184F" = "darkgreen", "MDR1-N86Y" = "darkred")) +
  scale_fill_identity(name = "Legend", guide = "legend") +
  labs(x = "Year", y = "Posterior Prevalence",
       title = "Posterior Prevalence of PfMDR1 Polymorphisms Over Time",
       subtitle = "Shade is 95% credible interval",
       fill = "95% Credible Interval") +
  theme_minimal()


```
**seeing if the code works**
**test MDR 1246, 2008**
```{r}
#in our original posterior df Mut_N = 22, N = 40
# in our prior information WT_N = 35, Mut_N = 111

percolateR(Mut_N_Prior = 35, WT_N_Prior = 111, Mut_N_New = 22, N_New = 40)

```

